Internet of Things: a Real-World Deployment Towards Smart Grids Integration

### 2013-07-05 trunk/doc/thesis/report.tex
major correction done. Not yet done

\documentclass[oneside,12pt,a4paper,final]{book}
%\usepackage[margin=3cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{arabtex}
\usepackage{utf8}
\newcommand{\HRule}{\rule{\linewidth}{0.6mm}}
%\usepackage{hyperref}
\usepackage{listings}
\lstset{language=C}
\usepackage[acronym]{glossaries} 
\makeglossaries
\author{Nacer Khalil}
\title{Internet of Things: a Real-World Deployment Towards Smart Grids Integration}

@@

@@

\chapter{Abstract}
\paragraph{}
The Information Technology is advancing exponentially and largely bigger impacting our daily lives. The current evolution in computing is called the \gls{iot}. It is smoothly evolving from an Internet of people to and Internet of Things. By 2020, it is expected to have 50 billion "Things" connected to the Internet. 
\paragraph{}
However such a migration induces a strong level of complexity when handling interoperability between heterogeneous things, i.e. Wireless Sensors, RFID, and mobile computing.
\paragraph{}
In this thesis, we focus on the  integration of wireless sensors into \gls{iot}. This prologues for a future integration of \gls{sg}\glsreset{sg} into \gls{iot}. In this work, we deployed a smart building testbed for energy optimization.
\paragraph{}
The system is using different technologies such as \gls{tinyos}, \gls{nesc}, \gls{j2se} and  \gls{http}.

@@

\section{Importance of Study}
\paragraph{}
The global electrical grid is one the bases that gives us this quality of life that most people around the world enjoy. The global electrical grid is sometimes called the largest machine ever made by man. \cite{ref1}. It a widespread machine, distributed but an interconnected system that provides power to households, companies and other organisations and therefore the world's economy is counting on it. The electrical grid knows numerous problems such as blackouts, current instabilities that cost billions of dollars yearly to both the utilities and insurances. The grid suffers from the problem of peak hours where energy consumption is high and in most cases, one single utility cannot provide.
\paragraph{}
The \gls{sg} comes as a solution to the problems that suffers from the traditional electrical grid. \gls{sg} is an intelligent, auto-balancing, and self-monitoring electrical grid \cite{ref2} and whose energy sources are distributed between utilities and consumers. The nature of the \gls{sg} as having distributed energy sources reduces numerous losses that arise from long energy transmission lines and also from the overload of the electrical lines.
\paragraph{}
Another revolution began and it is taking place in the information technology and is called \glsreset{iot}\gls{iot}. \gls{iot} is the era of autonomous machine able to produce and consume information and being able to act autonomously, without the guidance of humans, to performs their tasks. Cisco predicts that the number of devices to be connected to the Internet in 2020 will break the 50 billions devices \cite{ref3} whereas the world population will be around 7.6 billion. That makes 6.58 devices connected for each person in the world whereas in 2010, the number of devices connected was 12.5 billion. This clearly shows that the shift to the \gls{iot} has already started.
\paragraph{}
\glsreset{sg}\gls{sg} can and will benefit from the \gls{iot}. The integration of the \gls{sg} into the \gls{iot} makes this study and thesis of high importance in order to solve as much as possible the problems of the traditional grid but at the same bring a new dimension to the power grid by benefiting from the numerous possibilities offered by the \gls{iot}.

\section{Rationale of Study}
\paragraph{}
This project has been done in order to investigate ways the \gls{iot} can benefit the \gls{sg}. The Smart Grid as a set of philosophical ideas counts on a set of novel functionalities. One of these ideas is the two-way communication that will make possible the utility to communicate with consumers and vice versa. This concept is possible with the Internet Of Things as devices in \gls{iot} communicate between them by producing and consuming information. Numerous ideas are based on this two-way communication such as the \gls{dr} which offers a set of services for both the power provider and the consumer.
\paragraph{}
The \glsreset{iot}\gls{iot} counts on different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and \gls{6lowpan}\glsreset{6lowpan} which make the Internet of Things a very heterogeneous and complex system \cite{ref4}. One of the challenges in \gls{iot} is to deal with such heterogeneity and take the most out of it in order to make the communication as smooth and transparent as possible. This heterogeneity should be made transparent to any two parties trying to communicate in \gls{iot}.

\section{Problem Statement}
\paragraph{}
The \glsreset{sg} \gls{sg} counts on the two-way communication to build to a whole communication framework and a protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of this last. From the consumer's side or more precisely the household's side, there are smart home energy management systems in the market nowadays but none of them offers an integration with \gls{sg}. To have such an integration with \gls{sg}, there is a need to first build this two-way communication, to make it possible for any appliance to report its real-time energy consumption, and to be able to be controlled remotely with this infrastructure. Such pillar is missing for the time being, and therefore the integration of the smart home within the smart grid is not possible. Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two directions.
\paragraph{}
\gls{sg} counts on various technologies to achieve its main goals. This usage of different technologies brings another problem which is the heterogeneous nature of the system. As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system. As a matter of fact, a software should be built within the system, and would constitute an important part of it, whose main goal is to take care of making the communication technology-transparent and give the possibility to all components of the system to work hand-in-hand in order to achieve their common goals independently of the data link layer (Ethernet, Wifi or Zigbee) or the network layer technology (\gls{ipv4} or \gls{ipv6}) they are using.

\section{Purpose of Study}
 \paragraph{}
The goal of the study is to solve the problems introduced in the problem statement section. This means creating the two-way communication, and solving the heterogeneity that arises from integrating smart homes with the smart grid and making each components in the smart home as part of the Internet of Things by making it communicate with any device in the Internet by having an \gls{ipv6} address that is built by making use of \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and with the least computational resources.
\paragraph{}
The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on this possibility. To do it, there was a need to first introduce \gls{6lowpan} within every device in the \gls{wsn}. This would make every sensor node also known as mote identified and ready to communicate and be part of the \gls{iot}. This communication will go in the two directions and based on that, the two-way communication will be built.
\paragraph{}
The second purpose of this study is to solve the heterogeneity problem that arises with the integration of smart homes within the smart grid and also when making the whole system as part of the \glsreset{iot}\gls{iot}. Solving such a problem means having an additional component that will track all communication packets and change them to fit the needs of the destination component. This will make of this additional component a middleman and therefore make the heterogeneity problem disappear.

\section{Scope of Study}
\paragraph{}
The thesis aims to solves the two problems presented in section 1.3. To do so, a variety of tools and technologies were used for this purpose.The two problems were not solved one independently of the other but even though the problems seem separate, one cannot solve without having to solve the other. In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which supported \gls{ipv6} over \gls{6lowpan}, other networks only supported \gls{ipv4} and therefore could not communicate, and this was due to the heterogeneity problem. This is why the two-way communication was created but only after solving the heterogeneity problem, everything worked.
The system is divided into four subsystems:
\begin{itemize}
\item The fully \gls{ipv6} supported \gls{wsn}
\item The \gls{tinyos} program for sensor transmission and appliance request handler
\item The heterogeneity handler and middleman gateway server
\item The mobile client
\end{itemize}
\paragraph{}
Each of these four component is there to solve part of the problem. Each of these subsystems will be discussed extensively in the Chapter 4 from the architectural aspect of the component, and in Chapter 5 from the deployment point of view. 
\paragraph{}
It is to be noted that some of the concepts will be discussed from the architectural point in Chapter 4 but not from an implementation aspect in Chapter 5. The sole reason behind is the fact that some of the equipment required for this thesis did not arrive on time and therefore the parts in the implementation requiring such equipment were omitted to meet the deadlines.
\section{Outcome of the Study}
\paragraph{}
The outcome of the thesis is to design and implement a test-bed that is able to show the two-way communication in action. It will also have to show concepts of the \glsreset{iot}\gls{iot} which are that every node in the system can communicate autonomously and with any other node in the Internet. Another \gls{iot} concept is the ability of any node in the \gls{iot} to consume and produce information that is communicated through the network.
\paragraph{}
The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors that are attached to appliance in order to sense their consumption and this information is produced and transmitted periodically. The consumption of data by is shown by external nodes in the \gls{iot} sending requests to the sensor node in the \gls{wsn} in order to control the appliance's status i.e. turn On the television. This will prove the consumption of information by the nodes in the \gls{iot}.
\paragraph{}
The test-bed should also show that communication between any two nodes in the \gls{iot} is successful, robust and reliable and that the heterogeneity of the system is made transparent to the nodes in the system by the creation of the software tool residing in the gateway which is between the \gls{iot} and the outside world. This software tool is seen as a heterogeneity handler and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
\paragraph{}
The testbed will able to show the two-way communication in action and will also show that the heterogeneity does not form a barrier by its successful handling.

@@

The thesis is is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 

@@

It is an introductory chapter that gives a flavor of the project that is to be studied in this thesis and also introduces the problem and the solution proposed in this topic. This chapter also puts the reader into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}. 

@@

It is dedicated to the \glsreset{iot}\gls{iot}. It presents the topic, talks about the importance of \gls{iot}, its impact on the world, the actual Internet, the future of Internet and also how will \gls{sg} benefit from it.

@@

It will present the architecture of the system and the test-bed. It will discuss the system from different aspects, and how were the two problems solved from the architectural aspect.

@@

This chapter will mostly discuss the implementation system. It will start by presenting the hardware used in this project, then move to the technology enablers. Each component of the system will be discussed extensively and move to the integration of the components to form the test-bed.

@@

It will start first by presenting the findings then move to explain the experiment that was done to measure the performance of the system. The data gathering section will show how and what data was gathered and then move to analyzing and interpreting the data.

@@

These are the last two chapters of the thesis that will conclude the thesis, provide directions about future work and recommendations to the reader what parts of the system could be ameliorated and how it can be achieved.

@@

The actual electrical grid is based on a technology that is at least one century old \cite{ref5}. As a result, there is a need to change as to meet to 21st century requirement such as having more green energy in order to reduce the pollution caused mainly by the use of fossil energy. As a result, research has introduced a new technology called \glsreset{sg}\gls{sg}. Although it is still a research project and numerous issues related to \gls{sg} are left unanswered, nations across the globe started to deploy \gls{sg} project. It is also noted that in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.

@@

The smart grid brings around new concepts that will achieve efficiency, reliable and security to the power grid. The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6} and this reduce the losses of energy in the transmission.  This distribution by itself enables the power grid this to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer but households can now produce and even sell the extra power. This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producer and sell the excess power to the grid \cite{ref7}. This will allow the power grid to provide power to consumer in a "free market" fashion which means that it will obey to the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}

@@

The \glsreset{iot}\gls{iot} has known high acceptance by the players in the Internet. Nowadays, it is estimated that 12.5 billion devices are connected to the Internet and that this trend will continue to read 50 billion connected devices by the end of the decade \cite{ref3}. \gls{iot} brings several benefits to the actual Internet and these benefits will pull \gls{sg} to be part of it. Consequently, research has taken this direction which is to integrate the \gls{sg} within the \gls{iot} as most aspects of \gls{sg} will be enhanced by \gls{iot} \cite{ref8}.

@@

\gls{iot} counts on a set of technologies which are \glsreset{rfid}\gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power lines, WI-FI, Zigbee and other technologies \cite{ref8}. The smart grid's architecture based on \gls{iot} will be divided into three layers:

@@

Some of the research is discussing the need to replace Zigbee by Wi-Fi as the Wi-Fi offers higher bandwidth, non-line transmission ability, large-scale data collection and is highly cost-effective \cite{ref10} but still Zigbee has the sole advantage of very low energy consumption which is something required in \gls{sg} and \gls{iot}, because in systems that are installed far away from any maintenance team such as smart grid monitoring and self-healing devices, one cannot manage to move for the sole purpose of providing energy often. 

@@

From the architectural point of view, integrating \gls{sg} within \gls{iot} means having to address heterogeneity issues. An \gls{iot} gateway system that is based on Zigbee and GPRS protocols \cite{ref11}. This enables then to deal partly with the heterogeneity problem and therefore enable the \gls{wsn} to communicate with the mobile telecommunication network. Another solution to the heterogeneity problem is proposed where the solution is to present a new, light-weight web service transport protocol called Lean Transport Protocol (LTP) that will allow transparent exchange of web service messages between all kinds of devices \cite{ref12}. This protocol is platform-independent, low energy consumer. Some researchers claim that the heterogeneity is mainly coming from the fact that there are different \gls{wsn} devices that do not necessarily use the same standards and protocols. They propose to move all the \glspl{wsn} to all-IP networks as this would remove most of the heterogeneity but there are already deployed legacy \glspl{wsn} \cite{ref13}. They sketch an architecture capable of converting all the \glspl{wsn}, new and legacy, to support \gls{ipv6}.\glsreset{sg}

\chapter{Internet Of Things}
\section{The actual Internet}
\paragraph{}
The \glsreset{iot}\gls{iot} refers to the evolution of the actual Internet.The term \gls{iot} was used for the first time by Kevin Ashton in 1999 \cite{ref17}. The human beings will represent a minority in this network. \gls{iot} will become the biggest machine ever created by human beings. In the actual internet nowadays
\section{Introduction to \gls{iot}}
\paragraph{}
The \glsreset{iot}\gls{iot} is a network mostly ruled my machines, from the most powerful supercomputer and mainframes to the smallest devices such as sensor nodes. All will have one common point, being all connected to the same network and able to communicate, produce and consume information. The \gls{iot} is changing the world of computing. New paradigms, architectures, frameworks and prototypes are being built everyday to support it \cite{ref15}. New technologies such as \glsreset{rfid}\gls{rfid}, \glsreset{m2m}\gls{m2m}, \glsreset{ipv6}\gls{ipv6}, \glsreset{6lowpan}\gls{6lowpan} and \gls{ims} have come to place to support the \gls{iot}. The whole world is now being connected including humans and objects \cite{ref18}.
\paragraph{}
The \glsreset{iot}\gls{iot} brings numerous advantages but numerous issues will have to solved before this network explodes to trillions of devices connected to it. The first issue is the interconnection of all the devices into the \gls{iot}. What hardware will support such an ultra high traffic network with an ultra high performance and availability requirements. Will ubiquitous computing or pervasive computing have architectures that deal with such infrastructure which will clearly form a bottleneck in this whole complex system.
\paragraph{}
There are three steps in the flow of event in the \gls{iot} \cite{ref17}.
\begin{enumerate}
\item The object must by identified. \gls{rfid} is used for this purpose in addition to the feature of the objects should be recognized and converted to electronic format
\item  The information must be transported to the information consumer.
\item The information should be processed intelligently.
\end{enumerate}

\section{Importance of \gls{iot}}
\paragraph{}
The actual Internet is evolving into the \glsreset{iot}\gls{iot} because it allows things that are not possible in the Internet we all know. Autonomy of devices, production and consumption by things other than humans is a need. New high level communication language such as \gls{owl} are being created and studied extensively to enable objects to communicate freely without the need to interact with humans to get directives and configurations.
\paragraph{}
The world economy is having more and more requirements from the Internet. Requirements such as the identification of every device in the world and its interconnection to the Internet. The ability to get real-time information from devices throughout the globe.
\paragraph{}

\section{The Benefits of \gls{iot}}
\paragraph{}
\gls{iot} brings numerous solutions to the new requirements. In the Internet of Things, objects will work and communicate freely. Any device will be able to send its sensory information and consume data according to its needs. From this simple idea arise a whole new set of applications that will benefit the world. Smart grids exploit this idea and make it part of its requirement \cite{ref8}. More and more applications are now being developed. Ideas such as smart farming, smart transportation, smart homes, smart health, smart postal and smart refrigerator are only a small set of applications that count on the evolution of Internet towards the Internet of Things \cite{ref19}.
\paragraph{}
The Internet of Things will be able to help the world community solve and minimize the impact of numerous problems it faces. With the deployment of applications as part of \gls{iot}, we can predict natural disasters by sending real-time sensory data about geological, volcanic, seismic, meteorological and spacial variables and help communities avoid loosing their lives and being affected by these natural disasters. We can monitor water scarcity on a real-time basis. We can monitor in real-time our health by incorporating microscopic sensors that will be part of the Internet of Things.

>>>>

\documentclass[oneside,12pt,a4paper,final]{book}
%\usepackage[margin=3cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{arabtex}
\usepackage{utf8}
\newcommand{\HRule}{\rule{\linewidth}{0.6mm}}
\usepackage[titletoc]{appendix}
%\usepackage{hyperref}
\usepackage{listings}
\lstset{language=C}
\usepackage[acronym]{glossaries} 
\makeglossaries
\author{Nacer Khalil}
\title{Internet of Things: a Real-World Deployment Towards Smart Grids Integration}

@@

\usepackage{etoolbox}
\makeatletter
\patchcmd{\thebibliography}{%
  \chapter*{\bibname}\@mkboth{\MakeUppercase\bibname}{\MakeUppercase\bibname}}{%
  \chapter{Index}}{}{}
\makeatother

@@

\chapter{Abstract}
\paragraph{}
Information Technology is advancing exponentially and largely impacting our daily lives. The current evolution in computing is the \gls{iot}. It is smoothly evolving from an Internet of people to and Internet of Things. By 2020, it is expected to have 50 billion "Things" connected to the Internet. 
\paragraph{}
However such a migration induces a strong level of complexity when handling interoperability between heterogeneous Things, i.g. Wireless Sensors, RFID, and mobile devices.
\paragraph{}
In this thesis, we focus on the  integration of wireless sensors into \gls{iot}. This is q prologue for the future integration of \gls{sg} into \gls{iot}. In this work, we deployed a smart building testbed for energy optimization, and for integration into \gls{sg}\glsreset{sg}.
\paragraph{}
The testbed built provides IPv6 gatewaying, remote control of wireless sensor nodes which is built on top of \gls{6lowpan}\glsreset{6lowpan}, and mobile access to the system.
\paragraph{}
The system is using different technologies such as \gls{tinyos}, \gls{nesc}, \gls{j2se} and  \gls{http}.

@@

\section{Importance of Study}
\paragraph{}
The global electrical grid is the bases that gives us this quality of life that most people around the world enjoy. The global electrical grid is sometimes called the largest machine ever made by man \cite{ref1}. It is a widespread and distributed machine that provides power to households, companies and organizations. Therefore the whole world's economy counts on it. The electrical grid knows numerous problems such as blackouts, and current instabilities that cost billions of dollars yearly to both the utilities and insurances. The grid suffers from the problem of peak hours where energy consumption is high and in most cases, one single utility cannot provide.
\paragraph{}
The \glsreset{sg}\gls{sg} comes as a solution to the problems of the traditional electrical grid. \gls{sg} is an intelligent, auto-balancing, self-monitoring, and self-healing electrical grid \cite{ref2} and whose energy sources are distributed between utilities and consumers. The nature of the \gls{sg} as having distributed energy sources reduces numerous losses that arise from the long distance transmission lines, and also from the stress and the overload of the electrical lines.
\paragraph{}
Another revolution began, and it is taking place in the Information Technology. It is called the \glsreset{iot}\gls{iot}. \gls{iot} is the era of autonomous machine being able to produce and consume information, and being able to act autonomously, without the guidance of humans, to performs their tasks. Cisco predicts that the number of devices to be connected to the Internet in 2020 will break the 50 billions devices from 500 million in 2003. That would make 6.58 devices connected for each person in the world whereas that same ration was 0.07 17 years earlier \cite{ref3}. This clearly shows that the shift to the \gls{iot} will be exponential.
\paragraph{}
\glsreset{sg}\gls{sg} should benefit from the \gls{iot}. The integration of the \gls{sg} into the \gls{iot} constitutes the main topic of this study and goes along with increasing interest recently raised in both academia and industry, on the inherent correspondence between \gls{sg} and \gls{iot}. This  correspondence stems mainly from the need to disseminate information between the \gls{sg} components in order to make optimal decisions about energy usage. This information dissemination could only be optimal when deploying ubiquitous sensoring and metering, and using the available network infrastructure, i.e. the Internet.

\section{Rationale of Study}
\paragraph{}
This project has been done in order to investigate ways the \gls{iot} can benefit the \gls{sg}. The Smart Grid needs a set of novel functionalities. One of these ideas is the two-way communication that will make it possible for the utility to communicate with consumers and vice versa. This is possible in the Internet Of Things as devices in \gls{iot} communicate by producing and consuming information. Numerous ideas assume the existence of the two-way communication such as the \gls{dr} which offers a set of services for both the power provider and the consumer.
\paragraph{}
The \glsreset{iot}\gls{iot} uses different technologies such as \gls{wsn}, \gls{rfid}, \gls{ipv6} and \gls{6lowpan}\glsreset{6lowpan} which makes the Internet of Things a very heterogeneous and complex system \cite{ref4}. One of the challenges in \gls{iot} is to deal with this heterogeneity in order to make the communication smooth and transparent to the parties involved. This heterogeneity should become invisible to any two devices willing to communicate in the \gls{iot}.

\section{Problem Statement}
\paragraph{}
The \glsreset{sg} \gls{sg} relies on the two-way communication to build a whole communication framework and a protocol stack namely the \glsreset{dr}\gls{dr} and other services built on top of this last. From the consumer's/household's side or more precisely the household's side,there are smart home energy management systems in the market nowadays, but none of them offers an integration with \gls{sg}. To have such an integration with \gls{sg}, there is a demand for the two-way communication, which is to make it possible for any appliance to report its real-time energy consumption, and to be able to be controlled remotely. Such pillar is missing for the time being, and therefore the integration of the smart home within the smart grid is not possible. Also, in order for \gls{sg} to be integrated into the \gls{iot}, all devices should be able to communicate in the two directions.
\paragraph{}
\gls{sg} counts on various technologies to achieve its main goals. This use of different technologies comes with a problem which is the heterogeneous nature of the system. As numerous components count on different technologies, there is a need to interface them in order to make this heterogeneity transparent to the all components of the system. As a matter of fact, a software should be built within the system, and would constitute an important part of it, whose main goal is to take care of making the communication technology-transparent and give the possibility to all components of the system to work hand-in-hand in order to achieve their goals independently of the data link layer (Ethernet, Wifi or Zigbee) or the network layer technology (\gls{ipv4} or \gls{ipv6}) they are using.

\section{Purpose of Study}
 \paragraph{}
The goal of the study is to solve the problems introduced in the problem statement section. This means creating the two-way communication, and solving the heterogeneity that arises from integrating smart homes with the smart grid and making each components in the smart home as part of the Internet of Things by making it communicate with any device in the Internet by having an \gls{ipv6} address that is built by making use of \gls{6lowpan} whose goal is to provide the TCP/IP to the smallest devices and with the least computational resources.
\paragraph{}
The first purpose of this study is to create a testbed having the two-way communication and also being able to make use of it in order to build richer applications based on it. To do so, \gls{6lowpan} had to be integrated within every device in the \gls{wsn}. This would make every sensor node (mote) identifiable in the network and ready to communicate and be part of the \gls{iot}. This communication will go in both directions and based on \gls{6lowpan}, the two-way communication will be built.
\paragraph{}
The second purpose of this study is to solve the heterogeneity problem that arises with the integration of smart homes into the grid and also when making the whole system part of the \glsreset{iot}\gls{iot}. Solving such a problem means having an extra component that will track all  packets and change them to fit the standards of the destination component. This will make of this additional component a middleman and therefore make the heterogeneity problem disappear.

\section{Scope of Study}
\paragraph{}
This study aims to solves the two problems presented in section 1.3. To do so, a variety of tools and technologies were used for this purpose.The two problems could not be solved one independently of the other even though the problems seem separate, one cannot solve  the a problem without having to solve the other. In other words, after the creation of the two-way communication in the \glsreset{wsn}\gls{wsn} which adopts \gls{ipv6} over \gls{6lowpan}, other networks only support \gls{ipv4} and therefore the communication is not possible, and this was due to the heterogeneous nature of the system. This is why the two-way communication was created, but only after solving the heterogeneity problem, everything worked.
The system is divided into four subsystems:
\begin{itemize}
\item The fully \gls{ipv6} supported \gls{wsn}
\item The heterogeneity handler and middleman gateway server
\item The middle-ware
\item The mobile client
\end{itemize}
\paragraph{}
Each of these four components is there to solve part of the problem. Each of these subsystems will be discussed extensively in the Chapter 4 from its architectural aspect of the component, and in Chapter 5 from its deployment point of view. 
\paragraph{}
It is to be noted that some of the components will be discussed from the architectural point in Chapter 4 but not from an implementation aspect in Chapter 5. The sole reason behind is the fact that some of the equipment required for this thesis were not delivered on time and therefore the parts in the implementation requiring such equipment were omitted to meet the deadlines.
\section{Outcome of the Study}
\paragraph{}
The outcome of the thesis is to design and implement a test-bed that is able to show the two-way communication in action. It will also have to show its integration into the \glsreset{iot}\gls{iot} by showing that every node in the system can communicate autonomously and with any other node in the Internet. Another \gls{iot} concept is the ability of any node in the \gls{wsn} to consume and produce information that is communicated through the network.
\paragraph{}
The autonomous production of information is shown by the fact that every sensor node in the \gls{wsn} reads sensory data from its sensors, that are attached to appliance, in order to sense their consumption and this information is produced and communicated periodically. The consumption of data is proved by showing the execution of the commands communicated through the network by external nodes in the \gls{iot}. These commands request the sensor node to turn On or Off an appliance.
\paragraph{}
The test-bed should also show that communication between any two nodes in the \gls{iot} is successful and reliable, and that the heterogeneity of the system is transparent to the nodes in the system and this is done by creating a software tool deployed in the gateway which is between the \gls{wsn} and the outside world. This software tool is perceived as a heterogeneity handler and also a communication middleman that translates the communication to the appropriate technology needed by the recipient.
\paragraph{}
The testbed will be able to show the two-way communication in action and will also show that the heterogeneity is not a barrier.

@@

The thesis is divided into eight chapters. Each of the chapter approaches thesis from a different aspect. 

@@

It is an introductory chapter that gives a flavor of the project that is  studied in this thesis and also introduces the reader to the problem and the solution proposed in this context. This chapter also puts into perspective by introducing some of the main concepts related to the \glsreset{sg}\gls{sg} and the \glsreset{iot}\gls{iot}. 

@@

It is dedicated to the \glsreset{iot}\gls{iot}. It presents the topic, talks about the importance of \gls{iot}, its impact on the world, the actual Internet and the future of Internet.

@@

It will present the architecture of the test-bed. It will discuss the system from different aspects, and how were the problems stated in section 1.3 solved from the architectural aspect.

@@

This chapter will mostly discuss the system deployment. It will start by presenting the hardware used in this project, then move to the technology enablers. Afterward, Each component of the system will be discussed extensively and move to the integration of the components to form the test-bed.

@@

It will start first by presenting the findings then move to explain the experiments that were conducted. Afterwards, the data gathered will be shown, analyzed and interpreted.

@@

These are concluding chapters of this study. They provide directions about future work, recommendations to the reader about what parts of the system could be ameliorated and how it can be achieved.

@@

The actual electrical grid is based on a technology that is at least one century old and its problems are becoming more and more costly \cite{ref5}. Therefore, there is a need to change it as to meet to 21st century requirements such as having more green energy in order to reduce pollution caused mainly by the use of fossil energy. As a matter of fact, research introduced us to new technology: the \glsreset{sg}\gls{sg}. Although it is still a research project and has not reached maturity yet, nations across the globe started to deploy \gls{sg} projects. The \gls{sg} is relying on Information Technology to solve many of the problem existing in the traditional grid. Consequently, in order for the grid to become more reliable, secure and efficient, a bi-directional information and communication infrastructure should be embedded within the grid \cite{ref5}.

@@

The smart grid brings new concepts that will make it achieve greater efficiency, reliability and security compared to the traditional grid. The first one is the introduction of the \gls{dg} by exploiting \gls{res} \cite{ref6} and this reduce the losses of energy in the transmission.  This distribution by itself enables the power grid to move from the one way energy distribution to a two-way energy distribution where the utility will not be the sole energy producer, but households can now produce and even sell the extra electricity. This is known nowadays as the feed-in-tariff which is a policy that allows the consumers to become producers and sell the excess power to the grid \cite{ref7}. This will allow the power grid to provide electricity to consumer in a "free market" fashion which means that it will obey to the laws of supply and demand \cite{ref7}.\glsreset{iot} \glsreset{sg}

@@

The \glsreset{iot}\gls{iot} has been highly accepted by the players in the Internet as its adoption is starting smoothly. Nowadays, it is estimated that 12.5 billion devices are connected to the Internet and that this trend will continue to reach 50 billion devices by the end of the decade \cite{ref3}. \gls{iot} brings several benefits to the actual Internet and will pull the \gls{sg} to make this part of it. Consequently, research has taken the direction to integrate the \gls{sg} within the \gls{iot} as most of its aspects favor  the \gls{sg} \cite{ref8}.

@@

\gls{iot} counts on a set of technologies which are \glsreset{rfid}\gls{rfid}, \gls{m2m} \cite{ref9}, public communication networks such as 2G/3G mobile networks, power line communication, WI-FI and Zigbee\cite{ref8}. The smart grid's architecture based on \gls{iot} will be divided into three layers:

@@

Research discusses the benefits to replace Zigbee by Wi-Fi as this last offers higher bandwidth, non-line transmission ability, large-scale data collection and is highly cost-effective \cite{ref10}. Still, Zigbee has the exclusive advantage of comsuming the least energy of the communication technologies that exist nowadays which is something required in \gls{sg} and \gls{iot}, especially for systems that are installed far away from any maintenance team such as smart grid monitoring and self-healing devices where one cannot manage to move for the sole purpose of charging these devices. 

@@

From the architectural point of view, integrating \gls{sg} within \gls{iot} means having to address heterogeneity issues. An \gls{iot} gateway system based on Zigbee and GPRS protocols helps to deal partly with the heterogeneity problem and therefore enable the \gls{wsn} to communicate with the mobile telecommunication network \cite{ref11}. Another solution to the heterogeneity problem is proposed with a new, light-weight web service transport protocol called Lean Transport Protocol (LTP) that will allow transparent exchange of web service messages between all kinds of devices. This protocol is platform-independent, low energy communication \cite{ref12}. Some researchers claim that the major source of heterogeneity the fact that there are different types of \gls{wsn} devices (Micaz, Mica2, Telosb...) that do not use the same standards and protocols. They propose to move the communication in \glspl{wsn} to an "all-IP" as this would remove most of the heterogeneity but as there are already deployed legacy \glspl{wsn}, the task is complicated to conduct. Fortunately, the researchers sketch an architecture capable of converting all the \glspl{wsn}, new and legacy, to support \gls{ipv6} \cite{ref13}.\glsreset{sg}

\chapter{Internet Of Things}
\section{The Actual Internet}
\paragraph{}
The \glsreset{iot}\gls{iot} refers to the evolution of the actual Internet.The term \gls{iot} was used for the first time by Kevin Ashton in 1999 \cite{ref17}. The human beings will represent in the future a very small minority in the Internet and as it will be rules by autonomous machines. \gls{iot} will become the biggest machine ever created by human beings as the smart grid will be part of it.
\section{Introduction to \gls{iot}}
\paragraph{}
The \glsreset{iot}\gls{iot} is a network mostly ruled my machines, from the most powerful supercomputer and mainframes to the smallest devices such as sensor nodes. All will have one common point, being all connected to the Internet and able to communicate, produce and consume information. The \gls{iot} is changing the world of computing. New paradigms, architectures, frameworks and prototypes are built and refined everyday to support it \cite{ref15}. New technologies such as \glsreset{rfid}\gls{rfid}, \glsreset{m2m}\gls{m2m}, \glsreset{ipv6}\gls{ipv6}, \glsreset{6lowpan}\gls{6lowpan} and \gls{ims} have come to place to establish the infrastructure for \gls{iot}. The whole world is now being connected including humans and objects \cite{ref18}.
\paragraph{}
The \glsreset{iot}\gls{iot} brings numerous advantages but many issues will have to solved before this network explodes to connecting trillions. The first issue is the interconnection of all the devices into the \gls{iot}. What hardware will support such an ultra high traffic network with an ultra high performance and availability requirements. Will ubiquitous computing or pervasive computing have architectures that deal with such infrastructure?
\paragraph{}
To better understand \gls{iot},The following are the three steps in its flow of information:
\begin{enumerate}
\item The object is identified by making use of \gls{rfid}. In addition, the objects should be recognized by sensing the environment and this information should be converted to electronic format
\item  The information is transported to the information consumer.
\item The information is processed intelligently \cite{ref17}.
\end{enumerate}

\section{Importance of \gls{iot}}
\paragraph{}
The actual Internet is evolving into the \glsreset{iot}\gls{iot} because it opens possibilities that are not in the actual Internet. The ability of Things to produce and consume information is a requirement. New high-level communication languages such as \gls{owl} were invented and studied extensively to enable objects to communicate freely without the need to interact with humans.
\paragraph{}
The world's requirements from the Internet are growing exponentially. Such new requirement are the accurate identification of Things, and the ability to get real-time information from devices throughout the globe.
\paragraph{}

\section{The Benefits of \gls{iot}}
\paragraph{}
\gls{iot} is armed with solutions to the new requirements. In the Internet of Things, objects will work and communicate freely. Any device will be able to send its sensory information, and consume data. From this simple idea arises a whole new set of applications that will benefit the world. Smart grids exploit this idea and make it part of its requirement in order to solve the problem in the electrical grid\cite{ref8}. More and more applications based on \gls{iot} are now being developed. Ideas such as smart farming, smart transportation, smart homes, smart health, smart postal and smart refrigerators are only a small subset of a myriad of applications that will come to life with \gls{iot} \cite{ref19}.
\paragraph{}
The Internet of Things will be able to help the world's community solve and minimize the impact of hazards it faces. With the deployment of applications which will be part of \gls{iot}, we can predict natural disasters by sending real-time sensory data about geological, volcanic, seismic, meteorological and spacial variables and help communities avoid losing their lives. We can also monitor water scarcity on a real-time basis.Finally, each one's health can be monitored in real-time by incorporating microscopic sensors that are able to communicate this Information through Internet.

\section{The Future of Internet}
\paragraph{}
The Internet counts billions of connected devices nowadays. This number is growing exponentially and this is not going to stop in the coming decades. It is estimated that there will be trillions of objects connected by the end of the century \cite{ref3}. In the next decade, we will watch a major change in the Internet. The network architectures we know nowadays will changes radically. The hardware architectures used now will evolve to meet the expectations of \gls{iot}. Applications based on \gls{iot} will explode and new possibilities will be open to the world community.
\paragraph{}
Privacy of information is going to be a serious issue in \gls{iot}. As every device will be monitored and controlled, privacy becomes an issue. The question is how will the Internet of Things keep its evolution without having to be affected by privacy and security arguments \cite{ref20}.
